Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-03-27 endofquarterpaper.qmd
more edits

n_material <- c(unique(signbase_full_clean$material))

@@

One factor that may influence the geographic structure of sites is the possibility that some sites functioned as aggregation sites. The aggregation/dispersion pattern is defined by @conkley_agg as a type of hunter-gather settlement pattern, in which groups both break off in dispersed fragments and congregate in large groups, typically for various subsistence or social reasons. Aggregation sites are the places where those fragments or individuals congregate, and some archaeological indicators of aggregation sites include proximity to certain environmental resources, as well as a more diverse artifact assemblage [@conkley_agg]. Previous studies have also identified evidence of aggregation/dispersion patterns among the early modern humans of the Upper Paleolithic [@Maher_Conkey_2019; @bahn1982inter; @Bourdier_2013; @Svoboda_2013; @bahn1982inter; @white1992rethinking; @Soffer_Adovasio_Enloe_Audouze_Zubrow_2010; @montet-white1994].

. It is therefore possible that the sites with multiple connections to both groups, such as La Ferrassie in the the Proto-Aurignacian, Grotte de la Verpillière I in the Early Aurignroacian, and Gargas in the Evolved Aurignacian acted as aggregation sites or served a similar purpose. However, without examining environmental data, this remains only a hypothesis at this time.

@@

@@

Our work adds to a growing awareness of diversity and interconnection within the Upper Paleolithic. Traditionally, the focus on the transition from Neanderthals to early modern humans is envisioned as a point of clear rupture with the Middle Paleolithic, thus leading some researchers to over-emphasize homogeneity [@Teyssandier_2008; @tey2006]. Previous studies have complicated this perception, investigating those patterns of diversity and heterogeneity within the Upper Paleolithic [@Belfer-Cohen_Goring-Morris_2002; @Maher_Conkey_2019; @Vanhaeren_d; @belfer-cohen2003; @bordes1970; @Shemer_Barzilai_Marder_2024; @Schwendler_2012; @bordes1963]. This study contributes to that awareness, suggesting permeable, yet distinct cultural variation and boundaries.

This study adds to our knowledge of social networks within the Aurignacian period, as our network analysis and perMANOVA test results suggest dynamic patterns of social exchange and connection. Previous works [@bahn1982inter; @Horiuchi_Takakura_2019; @Rogers_2013; @baker2024evidence; @Golovanova_Doronichev_Doronicheva_Sapega_Shackley_2021; @Gamble_1982] have used material evidence analysis to reconstruct Upper Paleolithic social networks, which research has pointed to as a potential driver of cultural and evolutionary growth [@Greenbaum_Friesem_Hovers_Feldman_Kolodny_2019; @Creanza_Kolodny_Feldman_2017; @Cullen_1996]. @Foley_Gamble_2009 places the emergence of inter-group complexity at around 30 k BP, however, our analyses show evidence of inter-connectivity beginning at around 41 k BP, perhaps suggesting an earlier emergence of inter-connectivity than previously predicted. We hope that dividing the data up into smaller time periods serves to rectify some of the issues in low temporal resolution that tend to affect results when analyzing older, more time-averaged data sets [@Gravel-Miguel_Coward_2023]. Additionally, by focusing on change over time, we are able to see social networks as fluid and dynamic, rather than static and bounded [@Maher_Conkey_2019], as well as the mobility associated with hunter-gatherers [@Whallon_2006; @Padilla-Iglesias_Bischoff_2024].

>>>>

@@

One factor that may influence the geographic structure of sites is the possibility that some sites functioned as aggregation sites. The aggregation/dispersion pattern is defined by @conkley_agg as a type of hunter-gather settlement pattern, in which groups both break off in dispersed fragments and congregate in large groups, typically for various subsistence or social reasons. Aggregation sites are the places where those fragments or individuals congregate, and some archaeological indicators of aggregation sites include proximity to certain environmental resources, as well as a more diverse artifact assemblage [@conkley_agg]. Previous studies have also identified evidence of aggregation/dispersion patterns among the early modern humans of the Upper Paleolithic [@Maher_Conkey_2019; @bahn1982inter; @Bourdier_2013; @Svoboda_2013; @bahn1982inter; @white1992rethinking; @Soffer_Adovasio_Enloe_Audouze_Zubrow_2010; @montet-white1994]. It is therefore possible that the sites with multiple connections to both groups, such as La Ferrassie in the the Proto-Aurignacian, Grotte de la Verpillière I in the Early Aurignroacian, and Gargas in the Evolved Aurignacian acted as aggregation sites or served a similar purpose. However, without examining environmental data, this remains only a hypothesis at this time.

@@

This study adds to our knowledge of social networks within the Aurignacian period, as our results suggest dynamic patterns of social exchange and connection. Previous works [@bahn1982inter; @Horiuchi_Takakura_2019; @Rogers_2013; @baker2024evidence; @Golovanova_Doronichev_Doronicheva_Sapega_Shackley_2021; @Gamble_1982] have used material evidence analysis to reconstruct Upper Paleolithic social networks, which research has pointed to as a potential driver of cultural and evolutionary growth [@Greenbaum_Friesem_Hovers_Feldman_Kolodny_2019; @Creanza_Kolodny_Feldman_2017; @Cullen_1996]. @Foley_Gamble_2009 places the emergence of complex inter-group connectivity at around 30 k BP, however, our network analysis shows evidence of inter-connectivity beginning at around 41 k BP, perhaps suggesting an earlier emergence of complex inter-connectivity than previously predicted. We hope that dividing the data up into smaller time periods serves to rectify the low temporal resolution that tends to affect network analysis results when analyzing older, more time-averaged data sets [@Gravel-Miguel_Coward_2023]. Additionally, by focusing on change over time, we are able to see social networks as fluid and dynamic, rather than static and bounded [@Maher_Conkey_2019], as well as the mobility associated with hunter-gatherers [@Whallon_2006; @Padilla-Iglesias_Bischoff_2024].

@@

Our work also adds to a growing awareness of diversity within the Upper Paleolithic. Traditionally, the focus on the transition from Neanderthals to early modern humans is envisioned as a point of clear rupture with the Middle Paleolithic, thus leading some researchers to over-emphasize homogeneity [@Teyssandier_2008; @tey2006] and de-emphasize diversity. Previous studies have complicated this perception, investigating those patterns of diversity and heterogeneity within the Upper Paleolithic [@Belfer-Cohen_Goring-Morris_2002; @Maher_Conkey_2019; @Vanhaeren_d; @belfer-cohen2003; @bordes1970; @Shemer_Barzilai_Marder_2024; @Schwendler_2012; @bordes1963]. This study contributes to that awareness, suggesting permeable, yet distinct cultural variation and boundaries.
